Human Pose Estimation for Video Game Control

### 2022-05-08 paper/paper.tex
Added feature tracking for dino game

\subsubsection{Feature Classification For Jumping/Ducking}

>>>>

\subsubsection{Feature Classification For Jumping/Ducking} 
The second example video game we demonstrate our human-computer interface system 
on is the Google Chrome Dinosaur Game. The game accepts jumping and ducking commands 
in the form of up arrow key and down arrow key; however, our classifier is a 
little more complex than for the first example video game, as ducking and jumping 
require motion tracking over video to properly classify from human motion. That is, 
you cannot determine from a single frame of video whether or not the user is 
ducking or jumping.\\

In order to deisgn an appropriate classifier for this task, we first establish 
a subset of the keypoints as a more descriptive feature to track. In the case of 
the car steering example, this was done by reducing the set of keypoints to simply 
the left wrist and right wrist. For classification of jumping and ducking, this is 
accomplished by computing a "mean face" or the midpoint of a number of selected 
keypoints. For this application, we select the two eyes, two ears, and the nose, 
and compute the midpoint of these keypoints as our augmented feature to track in 
order to classify whether a user is jumping or ducking. The motivation for establishing 
this feature is to improve the robustness of our classification. Were we to simply 
track a single keypoint such as the nose of the user, our system would not be 
stable against random errors when the nose may not be properly classified or 
may not be classified at all from the phase 1 human pose estimation. Running 
relatively small versions of PoseNet, these types of errors are fairly common. 
Establishing the midpoint of the face as a feature to track allows us to track 
an equally useful but more robust feature.\\

After establishing the midpoint feature, we obtain graphs of the vertical 
position and derivative of vertical position of the midpoint feature over time 
as a user performs jumping and ducking motions in order to develop a game-specific 
classifier for this task. A few methods we considered and our analysis of them 
are given below:

\begin{enumerate}
    \item \textbf{Thresholding on Position}: The most simple method we could 
        employ to classify the movement of the midpoint face into either jumping 
        or ducking would be to threshold on the vertical position in the frame 
        of the midpoint of the users's face. However, this system is not invariant 
        to small motions of the user between jumps and ducks. For example, if the 
        user moves closer to or farther from the camera, this method breaks down.
    \item \textbf{Threshold on Derivative}: One method we considered would be 
        to simply threshold on the derivative of the y-position to discriminate 
        between periods of slow/no motion, and those where the user rapidly moves 
        up or down as one would during a jump or a duck. However, as we can see from 
        the graphs of the derivative over time, both jumps and ducks have periods of 
        high derivative in both the positive and negative direction. i.e) thresholding 
        on the deriviatve is not enough to discriminate between a duck and a jump, although 
        it can discriminate between jump/duck and no command.
    \item \textbf{Template Matching}: From the graph of the derivative over time, we can 
        see that a jump and a duck have inverted profiles on the graph. If we compute 
        the correlation between a template graph for a jump/duck and the current 
        graph at all timesteps, we would be able to discriminate between a jump and 
        a duck effectively. However, this method cannot be used in real time. As we can 
        see from either graph, a jump/duck occupies a significant portion of time. Computing 
        the correlation between a template feature and the motion of the midpoint face 
        in real time would only result in classification at the end of a jump or duck, which 
        would introduct unwanted latency into the system.
    \item \textbf{Threshold on Derivative with extra constraint}: The 
        method we implement in our system is a modification of simply thresholding on 
        the derivative of the y position. We threshold on the vertical position, with 
        the added constraint that at any time $t$, for the sample at that time to be 
        classified as a jump or a duck, the previous $N$ samples of the 
        derivative must be below the threshold. This ensures that only the 
        first period of rapid change in the motion of the midpoint face is 
        captured as a command. $N$ and the threshold are tunable parameters, 
        and in our final system, we set the threshold = $10$ pixels/sec and 
        $N = 3$.
\end{enumerate}

The final solution we implement for our feature tracking/classification for 
the Chrome dino game is as follows:

\begin{enumerate}
    \item For every frame of video, compute $y_t$, the midpoint of all keypoints 
        existing in the frame from the set 
        \{left eye, right eye, left ear, right ear, nose\}.
    \item Compute $y'_t$, the difference of $y_t$ and $y_{t-1}$.
    \item If $y'_t > 10, y'_{t-1} < 10, y'_{t-2} < 10$ and $y'_{t-3} < 10$: 
        Classify current frame as a duck.
    \item If $y'_t < 10, y'_{t-1} < 10, y'_{t-2} < 10$ and $y'_{t-3} < 10$:
        Classify current frame as a jump.
    \item If none of the above conditions are satisfied, classify current frame as 
        no command.
\end{enumerate}

\textit{Note that $y' > 10$ corresponds to a duck and $y' < 10$ corresponds 
to a jump as the image coordinates are zeroed at the top left of the frame.}
